---
title: "Analysis of an E-Commerce Sales Data"
output: html_notebook
---

#### Importing all the necessary libraries

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(scales)
```

#### Importing the data.

```{r}
data<- read.csv("Amazon_Sale_Report.csv")
```

#### Checking the column attributes of the dataset.

```{r}
colnames(data)
```

#### Covert into the appropriate naming convention(Sales.Channel *into* sales_channel)

```{r}
data<- clean_names(data)
colnames(data)
```

#### Checking the datatype of each column.

```{r}
str(data)
```

From this we can see that multiple column has wrong datatypes, such as Date is in the datatype chr instead of Date datatype.

```{r}
data$date<- mdy(data$date)
data$amount <- as.numeric(data$amount)
str(data)
```

We can see we have changed the datatype of date from chr to Date.

#### We can see that the column ship_state has some inaccuracy, like we can see that bihar has multiple entries like Bihar, BIHAR etc. We should correct it to get an explicit data analysis.

```{r}
data$ship_state <- tolower(data$ship_state)
data$ship_state <- str_to_title(data$ship_state)
```

This would alter the entries such that the first letter is uppercase and the rest are lowercase.

#### Now we check for the null values and we alter the values, or remove the null values.

```{r}
sapply(data, function(x) sum(is.na(x)))
```

```{r}
data$amount[is.na(data$amount)] <- 0
data$`ship_city`[is.na(data$`ship_city`)] <- "Missing"
data <- data[!is.na(data$`ship_postal_code`),]
```

```{r}
sapply(data, function(x) sum(is.na(x)))
```

### [**Answering the question related to the data**]{.underline}

#### **What is the average sales per month?**

```{r}
data %>% 
  mutate(Month = format(as.Date(date, "%m-%d-%y"), "%m")) %>% 
  group_by(Month) %>%
  summarise(Avg_Sales = mean(amount))
```

```{r}
data$Month <- floor_date(as.Date(data$date, "%m-%d-%y"), unit = "month")
```

```{r}
data2 <- data %>% 
  group_by(Month) %>%
  summarize(AvgSales = mean(amount))
```

```{r}
# Line chart 
ggplot(data2, aes(x = Month, y = AvgSales)) +
  geom_line() + 
  scale_y_continuous(labels = label_number()) + theme_minimal()
```

```{r}
# Heatmap of sales by month and category
ggplot(data) +
  geom_tile(aes(x = Month, y = category, fill = category)) + theme_minimal()
```

#### **Which market (state) generated the most sales on average?**

```{r}
data %>% 
  group_by(`ship_state`) %>%
  summarise(Avg_Sales = mean(amount)) %>%
  arrange(desc(Avg_Sales))
```

```{r}
data %>%
  group_by(`ship_state`) %>%
  summarise(Avg_Sales = mean(amount)) %>%
  arrange(desc(Avg_Sales)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(ship_state, -Avg_Sales), y = Avg_Sales)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(x = "State", y = "Average Sales") +
  theme_minimal() +
  coord_flip()
```

#### **Which category of products sell best?**

```{r}
data2_summary <- data %>%
  group_by(category) %>%
  summarise(Total_Sales = sum(amount)) %>%
  arrange(desc(Total_Sales))
```

```{r}
# Create a horizontal bar chart
bar_chart <- ggplot(data2_summary, aes(x = reorder(category, -Total_Sales), y = Total_Sales)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(x = "Category", y = "Total Sales") +
  scale_y_continuous(labels = label_number()) +
  theme_minimal()

# Display the bar chart
print(bar_chart)
```

#### **When were the best- and worst-selling periods?**

```{r}
data %>%
  mutate(Month = format(as.Date(date, "%m-%d-%y"), "%m")) %>%
  group_by(Month) %>%
  summarise(Sales = sum(amount)) %>%
  ggplot(aes(Month, Sales, fill = Month)) + 
  geom_col()+
  labs(
    title = "Sales by Month",
    y = "Total Sales"  
  )+
  scale_y_continuous(labels = label_number()) + theme_minimal()
```

#### **What is the average order value (AOV) for each product category?**

```{r}
data %>%
  group_by(category) %>%
  summarise(AOV = mean(amount))
```

```{r}
#Visualize AOV
data %>% 
  group_by(category) %>%
  summarise(AOV = mean(amount)) %>%
  ggplot(aes(x=category, y=AOV, fill = category)) +
  geom_col() +
  labs(title="Average Order Value by Product Category")+theme_minimal()
```
